Rclean: A Tool for Writing Cleaner, More Transparent Code

### 2020-02-05 joss/paper.Rmd
Adds R pack citations

The package's main functions are `clean` and `keep`. When provided a
file path to a script and the name of a result (or a set of results),
`clean` analyzes the script's code and extracts the lines of code
required to produce the results. By default, code is formatted
following general best practices recommended by the Tidyverse via the
`styler` package. This code can then be passed to the `keep` function,
which can either write the code to disk or copy the code to the user's
clipboard (if no output file path is supplied) and the user can paste
the code into another location (e.g. a script editor).

@@

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see [@Lang2018PackageComprehension]. To get an idea of what
data provenance is, take a look at the `code_graph` function. The plot
that it generates is a graphical representation of the prospective
provenance generated for `Rclean` \@ref(fig:prov-graph).

@@

The software is currently hosted on Github, and we recommend using the
`devtools` library to install directly from the repository
(https://github.com/ROpenSci/Rclean). The package is open-source and
welcomes contributions. Please visit the repository page to report
issues, request features or provide other feedback.

@@

Although the workflow described is impactful, we see promise in
extending it to interface with other clean code and reproducibility
tools. One example is the `reprex` package, which provides a simple
API for sharing reproducible examples. Another possibility is to help
transition scripts to function, package and workflow creation and
refactoring via toolboxes like `drake`. `Rclean` could provide a
reliable way to extract parts of a larger script that would be piped
to a simplified reproducible example, in the case of `reprex`, or,
since it can isolate the code from inputs to one or more outputs, be
used to extract all of the components needed to write one or more
functions that would be a part of a package or workflow, as is the
goal of `drake`.

@@

>>>>

The package's main functions are `clean` and `keep`. When provided a
file path to a script and the name of a result (or a set of results),
`clean` analyzes the script's code and extracts the lines of code
required to produce the results. By default, code is formatted
following general best practices recommended by the Tidyverse via the
`styler` package [@R-styler]. This code can then be passed to the
`keep` function, which can either write the code to disk or copy the
code to the user's clipboard (if no output file path is supplied) and
the user can paste the code into another location (e.g. a script
editor).

@@

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see [@R-CodeDepends]. To get an idea of what data provenance
is, take a look at the `code_graph` function. The plot that it
generates is a graphical representation of the prospective provenance
generated for `Rclean` \@ref(fig:prov-graph).

@@

The software is currently hosted on Github, and we recommend using the
`devtools` library [@R-devtools] to install directly from the
repository (https://github.com/ROpenSci/Rclean). The package is
open-source and welcomes contributions. Please visit the repository
page to report issues, request features or provide other feedback.

@@

Although the workflow described is impactful, we see promise in
extending it to interface with other clean code and reproducibility
tools. One example is the `reprex` package, which provides a simple
API for sharing reproducible examples [@R-reprex]. Another possibility is to help
transition scripts to function, package and workflow creation and
refactoring via toolboxes like `drake` [@R-drake]. `Rclean` could
provide a reliable way to extract parts of a larger script that would
be piped to a simplified reproducible example, in the case of
`reprex`, or, since it can isolate the code from inputs to one or more
outputs, be used to extract all of the components needed to write one
or more functions that would be a part of a package or workflow, as is
the goal of `drake`.

@@

<!-- Use overleaf papers.bib + knitr::write_bib -->
